---
title: "Week 6 Day 2 Homework"
output:
  html_document:
    df_print: paged
---
```{r, message=FALSE, warning=FALSE}
library(stats)
library(janitor)
library(here)
library(tidyverse)
here::here()
```
# 1 MVP

## Question 1.
Load the data, `clean_names()` and explore it.


```{r}
phones <- read_csv(here("data/20190928-items.csv")) %>% clean_names()
```

```{r}
dim(phones)

```
The dataset contains 792 observations of 9 variables.

```{r}
str(phones)
```

```{r}
head(phones)
```

## Question 2.
Find the brand with the highest numbers of phones in the dataset.



```{r}
phones %>%
  
  ggplot() +
  aes(x = brand) %>% 
  geom_bar(colour = "black", fill = "darkolivegreen4")
  
```

**Samsung have the highest number of phones in the data set.**

## Question 3.
For your top brand, plot the distribution of phone ratings as a probability 
density, overlaying a fitted normal distribution. Do you think the normal 
distribution provides a good model of these ratings?

```{r}
samsung <- phones %>% 
  filter(brand == "Samsung") 
samsung
```


The mean rating is 3.5733 and the standard deviation in rating is 0.6847859;
```{r}
samsung_stats <- samsung %>% 
  summarise(num = n(),
            mean_rating = mean(rating),
            sd_rating = sd(rating))

samsung_stats
```

```{r}


samsung %>% 
  ggplot() +
  aes(x = rating) +
  geom_histogram(aes(y = ..density..), 
                 col = "white", 
                 fill = "darkolivegreen4",
                 bins = 25) +
  stat_function(fun = dnorm,
                args = list(mean = samsung_stats$mean_rating,
                            sd = samsung_stats$sd_rating),
                colour = "red")
  


```


**On the appearance of the graph alone, the data appears to be a reasonable fit to **
**the normal distribution.**

## Question 4.
We can be more quantitative in comparing the distribution of top brand ratings with 
a normal distribution. You calculated the mean() and sd() of your distribution in 
an earlier question. Use these values to calculate the proportions of ratings within 
one-, two- and three standard deviations of the mean. Compare these proportions 
with those you expect for a normal distribution.

```{r}
# 1 SD
samsung %>%
  dplyr::filter(rating >= samsung_stats$mean_rating - samsung_stats$sd_rating) %>%
  dplyr::filter(rating <= samsung_stats$mean_rating + samsung_stats$sd_rating) %>%
  summarise(prop_within_1sd = n() / nrow(samsung))
```
**The proportion of data within 1 standard deviation is 0.793.**

```{r}
# 2 SD
samsung %>%
  dplyr::filter(rating >= samsung_stats$mean_rating - 2*(samsung_stats$sd_rating)) %>%
  dplyr::filter(rating <= samsung_stats$mean_rating + 2*(samsung_stats$sd_rating)) %>%
  summarise(prop_within_2sd = n() / nrow(samsung))
```



**The proportion of data within 2 standard deviations is 0.919.**

```{r}
# 3 SD
samsung %>%
  dplyr::filter(rating >= samsung_stats$mean_rating - 3*(samsung_stats$sd_rating)) %>%
  dplyr::filter(rating <= samsung_stats$mean_rating + 3*(samsung_stats$sd_rating)) %>%
  summarise(prop_within_3sd = n() / nrow(samsung))
```

**The proportion of data within 3 standard deviations is 0.980.**

The "textbook" normal distribution has standard deviation boundaries as follows;

```{r}
100 * (pnorm(q = 1) - pnorm(q = -1)) # 1 SD
100 * (pnorm(q = 2) - pnorm(q = -2)) # 2 SD
100 * (pnorm(q = 3) - pnorm(q = -3)) # 3 SD

```

So to summarise;

|SD |Normal Dist. |Samsung Dist. |
|----|----|----|
| 1 | 0.683 | 0.793 |
| 2 | 0.954 | 0.919 |
| 3 | 0.997 | 0.980 |

**So overall our data follows a normal distribution within +/- 1 standard deviation,**
**but drifts marginally outside the definition of a normal distribution beyond this.**


# 2 Extension
Investigate the use of the `qqnorm()` plot in R to determine whether data is 
normally distributed. What is a ‘quantile-quantile plot’? What do we expect to 
see in a qqnorm() plot if data is nearly normally distributed?



```{r}
qqnorm(samsung$rating)
qqline(samsung$rating)
```

A quantile - quantile plot is a visual representation of the quantiles of a data 
set plotted against the quantiles of a normal distribution.  If the data sample in 
question is also normal then the expected plot would show a line of points in a 
straight diagonal line.

The data from the Samsung ratings shown on the plot above do not completely follow 
the straight line, thereby given further indication that the data is not totally 
normally distributed.

The disadvantage of QQ plots is that like the normal distribution overlaid on a 
histogram, they are subjective and dependent on interpretation.