---
title: "Clustering homework"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(cluster)
library(factoextra)
library(dendextend)
library(corrplot)
library(broom)
library(here)



here::here()
```


```{r}
customers <- read_csv(here("data/mall_customers.csv")) %>% clean_names()
```

```{r}
# head(customers)
# str(customers)
```
Within the customer data there are five columns.  `customer_id` is redundant
in terms of clustering, but `spending_score` and `annual_income` look more 
promising, perhaps in conjunction with `age` and `gender`.

Check for missing values;
```{r}
customers %>% 
  summarise(across(.cols = everything(), ~ sum(is.na(.x))))
```

There are no missing values.


* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 

```{r}
customers_trim <- customers %>% 
  select(-customer_id)

customers_scale <- customers_trim %>% 
  mutate(across(where(is.numeric), scale),
         gender = if_else(gender == "Male", 1, 0)) # need numeric value for kmeans

customers_scale
```

```{r}
cust_clust <- kmeans(customers_scale,
                         centers = 6, # no. clusters desired
                         nstart = 25) # from 25 random sets

cust_clust
  
```


```{r}
tidy(cust_clust, col.names = colnames(customers_scale))
```
```{r}
augment(cust_clust, customers_trim)
```
```{r}
glance(cust_clust)
```





What do the spending score/annual income data look like;
```{r}
customers_scale %>% 
  ggplot() +
  aes(x = spending_score_1_100, annual_income_k, colour = gender) +
  geom_point() +
  labs(title = "Scaled Annual Income v Spending Score",
       x = "Spending Score",
       y = "Annual Income") +
  theme_bw()
```
It looks like there are probably five easily identificable clusters here.

* Perform k-means clustering and chose a value of k.

### Elbow Method

```{r}
fviz_nbclust(customers_scale,
             kmeans,
             method = "wss",
             nstart = 25)
```
The elbow method looks like it would be opimal to have 4 or 6 clusters, despite
what's shown in the ggplot earlier.

### Silhouette coeff. method

```{r}
fviz_nbclust(customers_scale,
             kmeans,
             method = "silhouette",
             nstart = 25)
```
The silhouette coefficient recommeds an optimal 6 clusters.  

### Gap Statistic

```{r}
fviz_nbclust(customers_scale,
             kmeans,
             method = "gap_stat",
             nstart = 25,
             k.max = 10)
```
...and the gap statistic recommends two clusters!

* Visualise the clustering for your chosen value of k.


```{r}
max_k <- 20

k_clusters <- tibble(k = 1:max_k) %>% 
  mutate(kclust = map(k, ~kmeans(customers_scale, .x, nstart = 25)), 
         tidied = map(kclust, tidy),
         glanced = map(kclust, glance),
         augmented = map(kclust, augment, customers_trim))

k_clusters
```


Select k = 5;

```{r}
clusterings %>% 
  unnest(cols = c(augmented)) %>% 
  filter(k == 5) %>% 
  ggplot() +
  aes(x = annual_income_k, spending_score_1_100, 
      colour = .cluster,
      label = .rownames) +
  geom_point(aes(colour = .cluster)) +
  geom_text(hjust = 0, vjust = -0.5, size = 3)
```


* Do you think the clustering seems a good fit for this data?

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 






