---
title: "Homework - features and elements of multiple regression"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(GGally)
library(ggfortify)
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')

```

```{r}
here::here()
```



<hr>

# MVP

<br>

1. Load the `housing_prices.csv` data set and undertake an initial exploration of the data. You will find details on the data set on the relevant [Kaggle page](https://www.kaggle.com/camnugent/california-housing-prices)


```{r}
house_prices <- read_csv(here("week_10/day_2/data/housing_prices.csv")) %>% clean_names()
house_prices
```


<br>

2. We expect the `total_rooms` of houses to be strongly correlated with `total_bedrooms`. Use `ggpairs()` to investigate correlations between these two variables.

```{r}
house_prices %>% 
  select(c(total_rooms, total_bedrooms)) %>% 
  ggpairs()
```


`ggpairs` shows a strong positive correlation between `total_rooms` and
`total_bedrooms`, i.e. `total_bedrooms` increases as `total_rooms` increases.

<br> 

3. So, we do find significant correlations. Let's drop `total_bedrooms` from the dataset, and use only `total_rooms` going forward.

```{r}
house_prices_nobed <- house_prices %>% 
  select(-total_bedrooms)
```


<br>

4. We are interested in developing a regression model for the `median_house_value` of a house in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `median_house_value` and the predictors (this may take a while to run, don't worry, make coffee or something).
  
  
```{r}
house_prices_nobed %>% 
  ggpairs()
```
  This resulting plot is horrible to work with, but it looks like there may be 
  some positive correlation between `median_house_value` and `median_income`.

  ii. Perform further `ggplot` visualisations of any significant correlations you find.

```{r}
options(scipen = 999) # "turn off" scientific notation...

# The median income and median house value are in different orders of magnitude.
# According to Kaggle.com the median income is in multiples of $10000, so that
# can be applied here;

house_prices_nobed <- house_prices_nobed %>% 
  mutate(median_income = median_income * 10000) 
  
house_prices_nobed %>%   
  ggplot() +
  aes(x = median_income, y = median_house_value) +
  geom_point() +
  labs(title = "Median House Value v Median Income",
       x = "Median Income",
       y = "Median House Value")
```

The ggplot of `median_house_value` v `median_income` confirms that there appears 
to be a positive correlation effect between the two variables.



<br>

5. Shortly we may try a regression model to fit the categorical predictor `ocean_proximity`. Investigate the level of `ocean_proximity` predictors. How many dummy variables do you expect to get from it?



```{r}
house_prices_nobed %>% 
  distinct(ocean_proximity)
```

There are five distinct values for `ocean_proximity`, so I would expect there to
be four dummy variables (leaving one as the reference).
    
<br>
 
6. Start with simple linear regression. Regress `median_house_value` on `median_income` and check the regression diagnostics.

```{r}
house_model <- lm(median_house_value ~ median_income, data = house_prices_nobed)
house_model
```

So the best fit line for these variables would be;
$${median\_house\_value} = 3.999 * {median\_income} + 45457$$
<br> 

7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

```{r}
house_model_v2 <- lm(median_house_value ~ median_income + total_rooms, 
                     data = house_prices_nobed)
house_model_v2
```

```{r}
summary(house_model_v2)
```

The p-value and Pr(>|t|) are well below 0.05 and the R-squared figure is 
reasonable.  How do the diagnostics look;


```{r}
autoplot(house_model_v2)
```

The diagnostics show a mixed result.  The Residuals v Fitted line is close to 0, 
but the data points do not look like they are randomly scattered around 0.

The Normal quantile - quantile plot is on the line in the middle of the data set,
but diverges from the line at both ends.

The ideal outcome for the scale-location plot is a band of fixed width above the 
x-axis, but this is not the case here.

In summary, this is probably not the best model for this data set.

# Extension
    
<br>

8. Try adding an interaction between `log(medium_income)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?



<br>

9. Find and plot an appropriate visualisation to show the effect of this interaction



