---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(ggfortify)
library(here)
library(modelr)




here::here()
```

```{r}
houses <- read_csv(here("data/kc_house_data.csv"))
houses
```

# Question 1
Tidy up the data ready for regression

```{r}
head(houses)
```

```{r}
houses_trim <- houses %>% 
  select(-c(id, date, sqft_lot, sqft_living15, sqft_lot15, zipcode)) %>% 
  mutate(waterfront = as.logical(waterfront),
         yr_renovated = if_else(yr_renovated == 0, FALSE, TRUE))
```

`condition` and `grade` are categorical ordinal data, so should probably be 
transformed.

```{r}
houses_trim <- houses_trim %>% 
  mutate(grade = case_when(
    grade <= 3 ~ "poor",
    grade > 3 & grade <= 5 ~ "below average",
    grade > 5 & grade <= 8 ~ "average",
    grade > 8 & grade <= 10 ~ "above average",
    grade > 10  ~ "excellent")) %>% 
  mutate(condition = case_when(
    condition == 1 ~ "poor",
    condition == 2 ~ "below average",
    condition == 3 ~ "average",
    condition == 4 ~ "above average",
    condition == 5  ~ "excellent"))
  
```


# Question 2
Check for aliased variables using the alias() function (this takes in a formula 
object and a data set).

```{r}
alias(lm(price ~ ., data = houses_trim))
```

The `alias()` function shows that `sqft_basement` can be calculated from
`sqft_living` and `not sqft_above`.  So `sqft_basement` and `sqft_above `could 
be dropped from the data set.

```{r}
houses_trim <- houses_trim %>% 
  select(-sqft_basement, -sqft_above)
```

# Question 3
Systematically build a regression model containing up to four main effects 
(remember, a main effect is just a single predictor with coefficient), testing 
the regression diagnostics as you go. 

```{r}
houses_tidy_numeric <- houses_trim %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- houses_trim %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- houses_trim$price

ggpairs(houses_tidy_numeric)
ggpairs(houses_tidy_nonnumeric)
```

At first glance it looks like there is a strong correlation between;

* price and sqft_living
* price and bathrooms

It also looks like `waterfront` has a strong effect on price.

## mod1a: price ~ sqft_living

```{r}
mod1a <- lm(price ~ sqft_living, data = houses_trim)
mod1a
```

So in isolation, $$price = -43580.7 + 280.6 * sqft\_living$$


```{r}
autoplot(mod1a)
```
This data does not look good.  None of the plots look acceptable, so sqft_living
can't be used as a predictor.


## mod1b: price ~ bathrooms

```{r}
mod1b <- lm(price ~ bathrooms, data = houses_trim)
mod1b
```

So in isolation, $$price = 10708 + 250327 * bathrooms$$

```{r}
autoplot(mod1b)
```
The diagnostics for bathrooms also look unacceptable!

## mod1c: price ~ waterfront

```{r}
mod1c <- lm(price ~ waterfront, data = houses_trim)
autoplot(mod1c)
```

...the Q-Q plot looks very non-linear!

I don't know where to go from here - I've tried four different predictors (I 
deleted one much earlier!) and they all look unusable in the diagnostic plots, 
which suggests to me that the basic assumptions of normality for linear 
regression are not being met.

I suspect that I'm doing something fundamentally wrong, maybe at the data
transformation stage, but I don't know what?
