---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

Question 1.

Read in the data and investigate it, e.g. variables, variable types, dimensions etc.

```{r}
wines <- read_csv("../../../de13_classnotes/week_01/day_4/4_extra_dplyr_practice/data/wine_ratings.csv")
```

```{r}
names(wines)
```

```{r}
head(wines, 10)

```

```{r}

dim(wines)

```

Question 2.

Select columns title, taster_name, points, price, and country, then drop all NAs. Save into a variable called wine_subset. Use this variable for all subsequent tasks.



```{r}

wine_subset <- wines %>% 
  select(title, taster_name, points, price, country) %>%
  drop_na()

```



Question 3.

Find the 5 countries with the highest mean points.

```{r}

wine_subset %>% 
  group_by(country) %>% 
  summarise(avg_points = mean(points)) %>% 
  slice_max(order_by = avg_points, n = 5)


```




Question 4.

Find the 5 countries with the lowest mean prices.

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(avg_price = mean(price)) %>% 
  slice_min(order_by = avg_price, n = 5)
```


Question 5.

Does the result differ if you use the median instead of the mean for Questions 3 and 4?



Question 6.

Find the names of all the tasters using distinct. Then take a random sample of 5 names using sample_n. Save these 5 names into a new variable called taster_sample which should be a character vector containing only the names.

Hint: You can use the function pull to extract a single column as a vector.


Question 7

Do the same thing for country, i.e. create a character vector called country_sample using distinct and pull, but this time use sample_frac to pick 10 countries at random.

Hint: What is the proportion of countries you want? (part / whole)


Question 8

Filter the rows of wine_subset so you are left only with data from the 5 tasters in your taster_sample and from the countries in your country_sample. Save this into a new variable called wine_sample.



Question 9

Using your wine_sample, create a new column difference_from_avg which contains the difference between each wine’s rating (points) and the average rating for that taster, then arrange the differences to see the highest positive difference first. [Hint: you will first have to calculate the average points for each taster]

