---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
revenue <- CodeClanData::qb_revenue_breakdown %>% janitor::clean_names()
monthly_sales <- CodeClanData::qb_monthly_sales %>% janitor::clean_names()
competitors <- CodeClanData::qb_competitors %>% janitor::clean_names()
device <- CodeClanData::qb_device_data %>% janitor::clean_names()
```

# Question 1.
Take the data in the dataset `qb_revenue_breakdown` and make a stacked bar chart 
showing the sources of revenue across the two years in the dataset.


Revenue has three columns, "product", "year" and "revenue";

```{r}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")

scale_fill_manual(col_scheme)
scale_color_manual(col_scheme)
```


```{r}
revenue %>% 
  ggplot() +
  aes(x = year, y = revenue, fill = product) + 
  geom_col(colour = "black") + 
  labs(title = "Sources of Revenue: 2018 - 2019") +
  theme_bw(base_size = 12)
```

# Question 2.
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit 
Unit 1’ - over the last year. This data is inside `qb_monthly_sales`, but you will 
need to use filter() or subset() (base R).

```{r}
kwikbit_sales <- monthly_sales %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  group_by(date) %>% 
  summarise(amount) 
  
ggplot(kwikbit_sales) +
  aes(x = date, y = amount) + 
  geom_line(colour = "#E89FE9") +
  geom_point(aes(x = date, y = amount), colour = "#E89FE9") +
  labs(title = "KwikBit Unit 1 Revenue by Month") +
  theme_bw(base_size = 12)
  
```

# Question 3.
Make a line chart showing monthly revenue and costs over the last year. 
This data is also in `qb_monthly_sales`

```{r}
operating_profit <- monthly_sales %>% 
  filter(date >= 2018-07-01 & (cashflow == "Cost" | cashflow == "Total Revenue")) %>% 
  group_by(cashflow, date) %>% 
  summarise(amount)
```

```{r}
operating_profit %>% 
  ggplot() +
  aes(x = date, y = amount, colour = cashflow) + 
  geom_line() +
  geom_point(aes(x = date, y = amount)) +
  labs(title = "12 Month Cost/Revenue") +
  theme_bw(base_size = 12)

```

# Question 4.
Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.


```{r}
competitors %>% 
  ggplot() +
  aes(x = year, y = revenue, fill = company) +
  geom_area() +
  scale_fill_brewer(palette = "Oranges") + 
  labs(title = "Fitness Tracker Sales by Company, 2015 - 2019") + 
  theme_bw(base_size = 12)

```

# Question 5.
Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
competitors %>% 
  ggplot() +
  aes(x = year, y = revenue, colour = company) +
  geom_line() +
  geom_point() +
  scale_fill_brewer(palette = "Set1") + 
  labs(title = "Fitness Tracker Sales by Company, 2015 - 2019") + 
  theme_bw(base_size = 12)

```

# Question 6.
Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in `qb_device_data`.


```{r}
device %>%
  ggplot() +
    aes(x = hours, y = counts, colour = device) + 
    geom_line() +
    facet_wrap(device ~ id) +
    theme_bw(base_size = 12)

  
  
```

