---
title: "Week 11 Weekend Homework | Text Mining"
output:
  html_document:
    df_print: paged
---

```{r}
library(janeaustenr)
library(tidyverse)
library(tidytext)
library(textdata)
```

# MVP 

Using the dataset austen_books() from the package janeaustenr:

1. Find the most common words in both Pride & Prejudice and Sense & Sensibility.

```{r}
# Change Pride & Prejudice into a tibble of words;

pride_book <- tibble(
  id = 1:length(prideprejudice),
  text = prideprejudice) %>% 
  unnest_tokens(word, text) 
pride_book %>% 
  count(word, sort = TRUE)


# and repeat for Sense & Sensibility

sense_book <- tibble(
  id = 1:length(sensesensibility),
  text = sensesensibility) %>% 
  unnest_tokens(word, text) 
sense_book %>% 
  count(word, sort = TRUE)


```


The most common words in "Pride and Prejudice" and "Sense and Sensibility" are 
stop words.


2. Find the most common words in both Pride & Prejudice and Sense & Sensibility, 
not including stop words.

Stop words can be removed by anti-joining the `stop_words` lexicon;
```{r}

pride_book %>% 
  anti_join(stop_words, by = "word") %>% 
  count(word, sort = TRUE)
```

```{r}
sense_book %>% 
  anti_join(stop_words, by = "word") %>% 
  count(word, sort = TRUE)
```



3. Find the most common sentiment words in both Pride & Prejudice and Sense & 
Sensibility.

Using the "bing" lexicon;

```{r}
pride_sentiment <- pride_book %>% 
  inner_join(get_sentiments("bing"), by = "word") %>%  # inner_join to avoid NAs
  count(word, sort = TRUE)
pride_sentiment
```

The top word is "miss", but this is a term of address rather than something having
been "missed", so can be removed;

```{r}
pride_sentiment <- pride_book %>%
  filter(word!= "miss") %>% 
  inner_join(get_sentiments("bing"), by = "word") %>%  # inner_join to avoid NAs
  count(word, sort = TRUE)
pride_sentiment
```


...and the same for "Sense and Sensibility"...

```{r}
sense_sentiment <- sense_book %>% 
  filter(word != "miss") %>% 
  inner_join(get_sentiments("bing"), by = "word") %>% 
  count(word, sort = TRUE)
sense_sentiment
```

# 2 Extension
Taking your results above. Can you create a plot which visualises the differences 
between the books?


Probably best to run a comparison of sentences, rather than words;


```{r}
book_pride <- tibble(
  text = prideprejudice,
  sentence = 1:length(prideprejudice)) %>% 
  unnest_tokens(word, text) %>% 
  anti_join(stop_words, by = "word") %>%   # remove stop words
  mutate(book = "pride")


book_sense <- tibble(
  text = sensesensibility,
  sentence = 1:length(sensesensibility)) %>% 
  unnest_tokens(word, text) %>% 
  anti_join(stop_words, by = "word") %>%  #remove stop words
  mutate(book = "sense")

sense_pride <- bind_rows(book_pride, book_sense) %>% 
  arrange(sentence) 

sentiment_scores <- sense_pride %>% 
  filter(word != "miss") %>%  
  inner_join(get_sentiments("afinn"), by = "word") %>% 
  # use afinn for numeric values
  group_by(book, sentence) %>% 
  summarise(n_words = n(), mean_sentiment = mean(value),
            total_sentiment = sum(value))


```

```{r}
sentiment_scores %>% 
  ggplot() +
  aes(x = sentence, y = mean_sentiment, colour = book) +
  geom_point(colour = "grey80", alpha = 0.1) +
  geom_smooth(method = "loess") +
  theme_light() +
  labs(title = "Mean Sentence Sentiment Comparison of 
       Pride and Prejudice v Sense and Sensibility",
       x = "Sentence No.",
       y = "Mean Sentiment Score (afinn)",
       colour = "Book")
```

The plot shows that both books score very similar results for sentiment on 
average, with Pride and Prejudice possibly having a happier ending than Sense
and Sensibility.  The mean remains above 0 for the duration of the books, so 
the good sentiments outweigh the bad.

Looking at total sentiment scores;

```{r}
sentiment_scores %>% 
  ggplot() +
  aes(x = sentence, y = total_sentiment, colour = book) +
  geom_point(colour = "grey80", alpha = 0.1) +
  geom_smooth(method = "loess") +
  theme_light() +
  labs(title = "Total Sentence Sentiment Comparison of 
       Pride and Prejudice v Sense and Sensibility",
       x = "Sentence No.",
       y = "Total Sentiment Score (afinn)",
       colour = "Book")
```

The total sentiment score between the books is closer, but once agani Pride and
Prejudice appears to finish on a slightly happier note.
