---
title: "Homework Week 6 Day 1"
output:
  html_document:
    df_print: paged
---


```{r, warning=FALSE, message=FALSE}
library(gtools)
library(prob)
library(tidyverse)
library(here)
library(janitor)

here::here()
```


```{r}

titanic <- read_csv(here("data/titanic.csv"))
titanic

```

```{r}
titanic %>% 
  distinct(class)
```

First create a new column to indicate who are crew and who are passengers;

```{r}
titanic <- titanic %>% 
  mutate(crew = class != "1st" & class != "2nd" & class != "3rd")

```

Determine the total number of crew:

```{r}
total_crew <- titanic %>%
  filter(crew == TRUE) %>%
  summarise(n())
total_crew
```
Thre were 890 crew.  How many survivors among them?

```{r}
total_crew_survived <- titanic %>%
  filter(crew == TRUE) %>% 
  filter(survived == "yes") %>%
  summarise(crew_surv = n())
total_crew_survived
```

So the proportion of survivors among the crew was:
```{r}
total_crew_survived / total_crew
```
 
 Similarly for passengers;
 
 Determine the total number of passengers:

```{r}
total_pasngr <- titanic %>%
  filter(crew != TRUE) %>%
  summarise(n())
total_pasngr
```
There were 1317 passengers.  How many survivors among them?

```{r}
total_pasngr_survived <- titanic %>%
  filter(crew != TRUE) %>% 
  filter(survived == "yes") %>%
  summarise(pasngr_surv = n())
total_pasngr_survived
```

So the proportion of survivors among the crew was:
```{r}
total_pasngr_survived / total_pasngr
```

So the proportion of survivors among the crew was 0.24.

The proportion of survivors among the passengers was 0.38.

**So passengers were more likely to have survived.**

# Question 1.
If I roll a six-sided dice once, what is the probability of getting a number less than 5?

There are six possible outcomes of a single dice roll;

```{r}
rolldie(1)
```
$$
probability(event) = (number(event outcomes) / number(all outcomes))
$$
The number of event outcomes if the number is to be < 5 is 4: {1, 2, 3, 4}

so the probability is
```{r}
4 / 6
```


# Question 2.
How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
rolldie(3)
```

**There are 216 possible outcomes.**


# Question 3.
In an experiment where we roll four dice:

 * how many outcomes are in the event where the total of the four dice equals 5, 6 
   or 7)?
 * what is the probability of this event?
 * what assumptions do we apply in answering part ii?


```{r}
roll_four <- rolldie(4) %>% 
  # add a column containing the sum of each row
  mutate(row_total = rowSums(.)) %>%   
  # filter for the desired outcomes
  filter(row_total == 5 | row_total == 6 | row_total == 7) %>% 
  # then count how many rows meet these requirements
  summarise(n())
roll_four
```

**There are 34 outcomes where the total of the four dice is 5, 6, or 7.**

The total number of possible outcomes is;

```{r}
rolldie(4) %>% 
  summarise(n())
```

**There are 1296 possible outcomes and we're interested in 34 of them, so the** 
**probability is;**

```{r}
34 / 1296
```
**Assuming we don't include any repeats.**


# Question 4.
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

```{r}
permutations(5, 5)
```

There are 120 permutations, of which only one is correct.  So the probability is;

```{r}
1 / 120
```

# Question 5.
I roll six dice:

 * what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the 
   dice, left to right?
 * what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the 
   dice regardless of order?

Hint
Have a look at the isin() function in the prob package. For example, to filter all rows containing 6, 1, 2 in that order, you could use isin(., c(6, 1, 2), ordered = TRUE)

```{r}
roll_six <- rolldie(6)
```

**Rolling six dice gives 46656 possible outcomes, so the probability of seeing**
**one specific outcome is;**
```{r}

options(scipen = 999) # turn off scientific notation...

1 / 46656
```
```{r}
dim(permutations(6, 6))
```
**There are 720 possible combinations of {1, 2, 3, 4, 5, 6}, so the probability of 
having any one of them is;**

```{r}
720 / 46656
```




# Question 6.
If I draw a single card from a 52-card deck, and it is black, what is the 
probability that its suit is spades?

In a 52-card deck, 26 cards are black.  Of those 26, 13 are Spades and 13 are 
Clubs.

**So if we pull a black card, the probability of it being a spade is;**
```{r}
13 / 26
```
Confirm in R;


```{r}
deck <- cards()
```

```{r}
deck <- deck %>%
  mutate(black_card = if_else(suit == "Club" | suit =="Spade",
                              as.logical(TRUE),
                              as.logical(FALSE)))
```

```{r}
deck_margin <- deck %>% 
  tabyl(suit, black_card) %>% 
  adorn_percentages(denominator = "all") %>% 
  adorn_totals(where = c("row", "col")) %>%    # note you have to adorn totals before title!
  adorn_title(col_name = "Is Black Suit")   # otherwise you get an error.
deck_margin
```
**The probability of being a spade in the whole pack;**

```{r}
deck %>% 
  summarise(prob = sum(suit == "Spade")/n())
```
**The probability of a black card being a spade;**

```{r}
deck %>% 
  filter(black_card == TRUE) %>% 
  summarise(prob = sum(suit == "Spade")/n())
```

