---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(infer)
library(here)

here::here()
```
# 1. MVP
## Task 1.
Load the data again, clean_names(), and re-familiarise yourself with it

```{r}
ames <- read_csv(here("data/ames.csv")) %>% clean_names()
```
```{r}
head(ames)
```
```{r}
glimpse(ames)
```

## Task 2.
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?

```{r}
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(colour = "white", fill = "darkolivegreen4", alpha = 0.8)
```
The lot_area data is right skewed as a result of a very small proportion of houses
having large lot areas.

## Task 3.
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.

```{r}
bootstrap_lot_200 <- ames %>% 
    rep_sample_n(size =200, replace = TRUE, reps = 5000) %>% 
    summarise(mean_lot_area = mean(lot_area))
```

```{r}
bootstrap_lot_200 %>% 
ggplot() +
  aes(x = mean_lot_area) +
  geom_histogram(colour = "white", fill = "darkolivegreen4", alpha = 0.8)
```


## Task 4.
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution

```{r}
ci_95 <- bootstrap_lot_200 %>% 
  get_confidence_interval(level = 0.95, type = "percentile") 
ci_95
```


```{r}
infer_resample <- ames %>% 
  # which variable are we interested in:
  specify(response = lot_area) %>% 
  # the replicates to resample:
  generate(reps = 5000, type = "bootstrap") %>% 
  # statistic we're interested in for each resample:
  calculate(stat = "mean")

infer_resample
```


```{r}
infer_ci_95 <- infer_resample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
infer_ci_95
```

```{r}
infer_resample %>% 
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95)
```




## Task 5.
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?

The 99% CI shading should be broader than 95% CI.

```{r}
infer_ci_99 <- infer_resample %>% 
  get_confidence_interval(level = 0.99, type = "percentile")
infer_ci_99
```

```{r}
infer_resample %>% 
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_99)
```


## Task 6.
Calculate the point estimate of the mean(lot_area)

```{r}
mean_infer <- infer_resample %>% 
  summarise(mean = mean(stat))
mean_infer
```
The Point Estimate of the mean is 10149.63.

# 2 Extension


## Task 1.
Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? [Investigate reps from 200 up to 50000, memory of your laptop permitting].
